---
title: "Recommendation Engines Final Project"
author: "Franz-Anton La Rosee"
date: "01/03/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Executive Summary:

## 1.1 Project Aim & Model Results

This markdown contains the development of two recommendation engines, where one provides recommendations for groceries based on past purchases and the other provides recommendations for movies, based on user ratings. Performance parameters for each model are as follows:

### Model Results - Recommendation Engine For Groceries

### Model Results - Recommendation Engine For Movies

Given the seperate nature, aim and construction of each recommendation engine, the model development section of this markdown is split into two parts, where section 2.1 contains the development for the groceries recommendation engine and section 2.2 contains the development for the movies recommendation engine. 

# 2. Model Development: Recommendation Engine For Groceries

## 2.1 Data Exploration

## 2.2 Model Pre-Selection

## 2.3 Data Preperation

## 2.4 Model Tuning

# 3. Model Development: Recommendation Engine For Movies

## 3.1 Data Exploration

## 3.2 Model Pre-Selection

## 3.3 Data Preperation

## 3.4 Model Tuning


```{r}

```




You can also embed plots, for example:

```{r, echo=FALSE}

```

